COVID19 - Forecast analysis

### 2020-03-12 draft_analysis_Italy.Rmd
Add home page, navigation and theme

```{r  plot data, echo=TRUE}
plot(dat_csv$data,dat_csv$totale_casi,ylab="Total Covid cases",xlab="Date")
mean(diff(log(dat_csv$totale_casi)))
```
The grow is exponential (rate of increase of about 25\%). 
We estimate the R0 parameter by means of a linear model. R0 indicates how contagious an infectious disease is. It's also referred to as the reproduction number.

@@

The slope b indicate the rate of exponetial increase, used for the R0 calculation.   
R code:  
fit1 <- lm(log(totale_casi)~t,data=dat_csv)

```{r data}
head(dat_csv)
fit1 <- lm(log(totale_casi)~t,data=dat_csv)
summary(fit1)
```
The prediction is  good ($R^2$ near 1).
However the tendency is for a reduction of the slope.

```{r model plot, echo=TRUE}
plot(dat_csv$t,log(dat_csv$totale_casi),ylab="log cases",xlab="time")
abline(coef(summary(fit1))[,1])
```
The slope coefficient estimated in the linear regression model can be used to estimate R0.

@@

The incubation period for the coronavirus is in mean 5.1 days with a range from 2-14 days.
Please se https://www.worldometers.info/coronavirus/coronavirus-incubation-period/.
In the calculation we considered an incubation period of 14 days for two reasons: 1) the majority are asymptomatic, contagiousness is greater than 5, maybe 14. A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 2) 14 days is the worst scenario to begin.
```{r r0 plot, echo=TRUE}
slope <-coef(summary(fit1))[2,1]; slope
slope.se <- coef(summary(fit1))[2,2]; slope.se
### R0 estimates and 95% IC 
R_0=slope*14+1;R_0
(slope+c(-1,1)*1.96*slope.se)*14+1
```
We want to make a short term forecast (14 days) with 3 scenario:  
-Scenario 1: 10 exposed people for each COVID-19 case and beta the same (no restrictions made or even no effects)
-Scenario 2: 5 exposed people for each COVID-19 case and beta reduced of 50% (-50% both exposed people and -50% COVID19 contagious power)
-Scenario 3: 3 exposed people for each COVID-19 case and beta reduced of 50%  (-70% both exposed people and -50% COVID19 contagious power)

@@

```{r initation plot, echo=TRUE}
# initial number of infectus
I0<-max(dat_csv$totale_casi); I0
# initial number of recovered
R0<-max(dat_csv$dimessi_guariti); R0
#beta 
beta0<-R_0/(14)
# italian poulation
N=60480000
# duration of COVID19 
duration<-14
#sigma0 is the coronavirus transmission rate fixed to 5%  (half of flu epidemic)
sigma0<-0.05
#mortality rate 
mu0<-1/(82*365.25) # 1/lifespan
```
We use the library(EpiDynamics)
```{r first scenario plot, echo=TRUE}
library(EpiDynamics)

# average number of single connections of an infected person
# less contacts, less probability of new infections
# we keep constant the other parameters
forecast<-14
parameters <- c(mu = mu0, beta = beta0, sigma = sigma0, gamma = 1/duration)
f1<-10
initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
seir1 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
f2<-5
initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
seir2 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
f3<-3
initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
seir3 <- SEIR(pars = parameters, init = initials, time = 0:forecast)

date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
plot(date,c(dat_csv$totale_casi,seir1$results$I[-1]*N),type="l",ylab="Number of infectus",xlab="time")
lines(date,c(dat_csv$totale_casi,seir2$results$I[-1]*N),col=2)
lines(date,c(dat_csv$totale_casi,seir3$results$I[-1]*N),col=3)
lines(date[1:dim(dat_csv)[1]],dat_csv$totale_casi,lwd=2)
legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
```

>>>>

```{r  plot data, echo=TRUE}
plot(dat_csv$data,dat_csv$totale_attualmente_positivi,ylab="Total Covid cases",xlab="Date")
mean(diff(log(dat_csv$totale_attualmente_positivi)))
```
The grow is exponential (rate of increase of about 25\%). 
We estimate the R0 parameter by means of a linear model. R0 indicates how contagious an infectious disease is. It's also referred to as the reproduction number.

@@

The slope b indicate the rate of exponetial increase, used for the R0 calculation.   
R code:  
fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)

```{r data}
head(dat_csv)
fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)
summary(fit1)
```
The prediction is  good ($R^2$ near 1).
However the tendency is for a reduction of the slope.

```{r model plot, echo=TRUE}
plot(dat_csv$t,log(dat_csv$totale_attualmente_positivi),ylab="log cases",xlab="time")
abline(coef(summary(fit1))[,1])
```
The slope coefficient estimated in the linear regression model can be used to estimate R0.

@@

The incubation period for the coronavirus is in mean 5.1 days with a range from 2-14 days.
Please se https://www.worldometers.info/coronavirus/coronavirus-incubation-period/.
In the calculation we considered an incubation period of 14 days for two reasons: 1) the majority are asymptomatic, contagiousness is greater than 5, maybe 14. A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 2) 14 days is the worst scenario to begin.
```{r r0 plot, echo=TRUE}
slope <-coef(summary(fit1))[2,1]; slope
slope.se <- coef(summary(fit1))[2,2]; slope.se
### R0 estimates and 95% IC 
R_0=slope*14+1;R_0
(slope+c(-1,1)*1.96*slope.se)*14+1
```
We want to make a short term forecast (14 days) with 3 scenario:  
-Scenario 1: 10 exposed people for each COVID-19 case and beta the same (no restrictions made or even no effects)
-Scenario 2: 10 exposed people for each COVID-19 case and beta reduced of 50% (-50% COVID19 contagious power)
-Scenario 3: 5 exposed people for each COVID-19 case and beta reduced of 50%  (-50% both exposed people and COVID19 contagious power)

@@

```{r initation plot, echo=TRUE}
# initial number of infectus
I0<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I0
# initial number of recovered
R0<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R0
#beta 
beta0<-R_0/(14)
# italian poulation
N=60480000
# duration of COVID19 
duration<-14
#sigma0 is the coronavirus transmission rate fixed to 5%  (half of flu epidemic)
sigma0<-0.05
#mortality rate 
mu0<-1/(82*365.25) # 1/lifespan
```
We use the library(EpiDynamics)
```{r  scenario plot, echo=TRUE}
library(EpiDynamics)

# average number of single connections of an infected person
# less contacts, less probability of new infections
# we keep constant the other parameters
forecast<-14
parameters <- c(mu = mu0, beta = beta0, sigma = sigma0, gamma = 1/duration)
f1<-10
initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
seir1 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
f2<-10
initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
seir2 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
f3<-5
initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
seir3 <- SEIR(pars = parameters, init = initials, time = 0:forecast)

date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
plot(date,c(dat_csv$totale_attualmente_positivi,seir1$results$I[-1]*N),type="l",ylab="Cases",xlab="time",main="Infected")
lines(date,c(dat_csv$totale_attualmente_positivi,seir2$results$I[-1]*N),col=2)
lines(date,c(dat_csv$totale_attualmente_positivi,seir3$results$I[-1]*N),col=3)
lines(date[1:dim(dat_csv)[1]],dat_csv$totale_attualmente_positivi,lwd=2)
legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
```
The 3 scenarios show how measures of restriction can help to reduce the number of infected.  
At the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (seir2$results$I*N)[length(seir2$results$I)]`).
In the next plot the cumulative number of infected.
```{r  cumulative plot, echo=TRUE}

plot(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir1$results$I[-1]*N)),type="l",ylab="Cases",xlab="time",main="Cumulative Infected")
lines(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir2$results$I[-1]*N)),col=2)
lines(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir3$results$I[-1]*N)),col=3)
lines(date[1:dim(dat_csv)[1]],cumsum(dat_csv$totale_attualmente_positivi),lwd=2)
legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
```

### 2020-03-12 draft_analysis_Veneto.Rmd
Add home page, navigation and theme

```{r setup, include=FALSE}
rm(list=ls())
###import italian dataset updated 10 March 2020
dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv",header=T)
dat_csv<-dat_csv[dat_csv$codice_regione==5,]
dat_csv$t<-1:dim(dat_csv)[1]
```

fit1 <- lm(log(totale_casi)~t,data=dat_csv)

```{r cars}
show(dat_csv)
fit1 <- lm(log(totale_casi)~t,data=dat_csv)
summary(fit1)
```

@@

```{r model plot, echo=TRUE}
plot(dat_csv$t,log(dat_csv$totale_casi),ylab="log cases",xlab="time")
abline(coef(summary(fit1))[,1])
```
The slope coefficient estimated in the linear regression model can be used to estimate R0, the number of 
```{r r0 plot, echo=TRUE}
slope <-coef(summary(fit1))[2,1]; slope
slope.se <- coef(summary(fit1))[2,2]; slope.se
### R0 estimates and 95%IC 
### I have used 14 for infection time, but it has a bimodal distribution (tested vs non tested)
R_0=slope*14+1;R_0
(slope+c(-1,1)*1.96*slope.se)*14+1
```
We want to make a short term forecast (14 days) with 3 scenario.
We fix a series of initial parameters.
```{r initation plot, echo=TRUE}
# initial number of infectus
I0<-max(dat_csv$totale_casi); I0
# initial number of recovered
R0<-max(dat_csv$dimessi_guariti); R0
#beta 
beta0<-R_0/(14)
# Veneto poulation
N=4800000
# duration of COVID19 
duration<-14
#sigma0 is the coronavirus transmission rate fixed to 5%  (half of flu epidemic)
sigma0<-0.05
#mortality rate 
mu0<-1/(82*365.25) # 1/lifespan
```
We use the library(EpiDynamics)
```{r first scenario plot, echo=TRUE}
library(EpiDynamics)
parameters <- c(mu = mu0, beta = beta0, sigma = sigma0, gamma = 1/duration)
# average number of single connections of an infected person
# less contacts, less probability of new infections
# we keep constant the other parameters
f1<-10
initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
seir1 <- SEIR(pars = parameters, init = initials, time = 0:14)

f2<-5
initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
seir2 <- SEIR(pars = parameters, init = initials, time = 0:14)

f3<-3
initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
seir3 <- SEIR(pars = parameters, init = initials, time = 0:14)

plot(c(dat_csv$totale_casi,seir1$results$I[-1]*N),type="l",ylab="Number of infectus",xlab="time")
lines(c(dat_csv$totale_casi,seir2$results$I[-1]*N),col=2)
lines(c(dat_csv$totale_casi,seir3$results$I[-1]*N),col=3)
legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
```

>>>>

```{r setup, include=FALSE}
rm(list=ls())
###import italian dataset updated 10 March 2020
dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv",header=T)
dat_csv<-dat_csv[dat_csv$codice_regione==5,]
dat_csv$t<-1:dim(dat_csv)[1]
plot(dat_csv$data,dat_csv$totale_attualmente_positivi,ylab="Infected",xlab="Date")
```

fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)

```{r cars}
head(dat_csv)
fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)
summary(fit1)
```

@@

```{r model plot, echo=TRUE}
plot(dat_csv$t,log(dat_csv$totale_attualmente_positivi),ylab="log cases",xlab="time")
abline(coef(summary(fit1))[,1])
```
The slope coefficient estimated in the linear regression model can be used to estimate R0, the number of 
```{r r0 plot, echo=TRUE}
slope <-coef(summary(fit1))[2,1]; slope
slope.se <- coef(summary(fit1))[2,2]; slope.se
### R0 estimates and 95%IC 
### I have used 14 for infection time, but it has a bimodal distribution (tested vs non tested)
R_0=slope*14+1;R_0
(slope+c(-1,1)*1.96*slope.se)*14+1
```
We want to make a short term forecast (14 days) with 3 scenario:  
-Scenario 1: 10 exposed people for each COVID-19 case and beta the same (no restrictions made or even no effects)
-Scenario 2: 10 exposed people for each COVID-19 case and beta reduced of 50% (-50% COVID19 contagious power)
-Scenario 3: 5 exposed people for each COVID-19 case and beta reduced of 50%  (-50% both exposed people and COVID19 contagious power)

We fix a series of initial parameters:
-I0: initial number of COVID-19 cases  
- R0: initial number of recovered  
- beta: the quantity connected to R0  
- N: Italian population   
- duration: infection duration of COVID-19   
- sigma0: the coronavirus transmission rate (half of flu epidemic)  
- mu0: the overall mortality rate  
```{r initation plot, echo=TRUE}
# initial number of infectus
I0<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I0
# initial number of recovered
R0<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R0
#beta 
beta0<-R_0/(14)
# Veneto population
N=4905000
# duration of COVID19 
duration<-14
#sigma0 is the coronavirus transmission rate fixed to 5%  (half of flu epidemic)
sigma0<-0.05
#mortality rate 
mu0<-1/(82*365.25) # 1/lifespan
```
We use the library(EpiDynamics)
```{r first scenario plot, echo=TRUE}
library(EpiDynamics)
forecast<-14
parameters <- c(mu = mu0, beta = beta0, sigma = sigma0, gamma = 1/duration)
f1<-10
initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
seir1 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
f2<-10
initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
seir2 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
f3<-5
initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
seir3 <- SEIR(pars = parameters, init = initials, time = 0:forecast)

date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
plot(date,c(dat_csv$totale_attualmente_positivi,seir1$results$I[-1]*N),type="l",ylab="Cases",xlab="time",main="Infected")
lines(date,c(dat_csv$totale_attualmente_positivi,seir2$results$I[-1]*N),col=2)
lines(date,c(dat_csv$totale_attualmente_positivi,seir3$results$I[-1]*N),col=3)
lines(date[1:dim(dat_csv)[1]],dat_csv$totale_attualmente_positivi,lwd=2)
legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
```
The 3 scenarios show how measures of restriction can help to reduce the number of infected.  
At the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (seir2$results$I*N)[length(seir2$results$I)]`).
In the next plot the cumulative number of infected.
```{r  cumulative plot, echo=TRUE}

plot(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir1$results$I[-1]*N)),type="l",ylab="Cases",xlab="time",main="Cumulative Infected")
lines(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir2$results$I[-1]*N)),col=2)
lines(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir3$results$I[-1]*N)),col=3)
lines(date[1:dim(dat_csv)[1]],cumsum(dat_csv$totale_attualmente_positivi),lwd=2)
legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
```
